### 2023-06-02 manuscript/manuscript.qmd
Added first two paragraphs to Introduction

---
title: Climate predictor selection and variable reduction methods influence MaxEnt model performance and predictions of climatic suitability 
author:
  - name: Clarke van Steenderen
    email: vsteenderen@gmail.com
    affiliations: 
        - id: some-tech
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: false
  - name: Guy F. Sutton
    email: g.sutton@ru.ac.za
    affiliations:
        - id: another-u
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: true
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Modelling
      formatting: preprint
      model: 3p
      layout: twocolumn
      cite-style: authoryear
      highlights:
        - Highlight 1
        - Highlight 2 
        - Highlight 3
---

# Introduction

Ecological models are important tools to aid the development and implementation of environmental policies and management programmes [@Addison2013; @Schuwirth2019; @Sutton2022]. These models are used for conservation planning [@Guisan2013], predicting the establishment and spread of invasive species [@Martin2020], implementing biological control programmes [@Sutton2019b; @Mukherjee2021], and forecasting species responses to environmental change [@Bocedi2014], amongst other applications. Species distribution models (SDM’s) are an example of ecological models that have become increasingly popular in recent years [@Elith2009]. SDM’s typically take the form of correlative or mechanistic models that correlate species presence/absences (or pseudo-absences) to environmental covariates to identify suitable climatic conditions for the study taxon [@Elith2011]. The Maximum Entropy species distribution model (hereafter ‘MaxEnt’) is amongst the most popular methods for climate modelling studies and has been shown to perform well compared to alternative modelling algorithms [@Wisz2008; @Phillips2017]. It uses maximum entropy to distinguish between environmental conditions where the focal taxon is present from environmental conditions at sites without confirmed presence records for the taxon [@Elith2011].

In recent years, a number of studies have investigated and demonstrated that computational choices made during the model building process can have a significant influence on resulting model outputs and inferences drawn [@Warren2011; @Webber2011; @Shcheglovitova2013; @Boria2017; @Sutton2022]. Despite its importance in the model building process, covariate selection methods have received considerably little attention to date [but see @Austin2011; @Fourcade2018; @Adde2023], whereby covariate selection refers to "identify[ing] the best subset of covariates out of a panel of many candidates, both from an ecological and statistical perspective [see @Adde2023 and references therein].